Time-domain multi-patch discontinuous Galerkin methods

### 2017-03-29 docs/IGA-DG/main.tex
update IGA docs (may be cruft now)

\newcommand{\half}{1/2}

@@

\date{}
\author{Jesse Chan, John Evans}
\title{B-spline DG notes}

@@

\begin{abstract}
Achieve extra $h$-resolution for the same CFL condition by using high order B-splines as a local approximation space on tensor product elements.

Achieve a spectral element-type implementation using quadrature nodes for B-splines. 
\end{abstract}

@@

\section{Stable timestep restrictions for high order DG methods}

We illustrate this using the advection equation.  
\begin{align*}
V^T A_h U &= \sum_{k} \LRp{\beta\cdot\Grad u,v}_{\Lk} + \LRa{f^*(u),v}_{\Ldk} \\
&\leq \nor{\beta}_{L^\infty}\sum_{k} \nor{\Grad u}\nor{v} + \nor{f^*(u)}_{\Ldk}\nor{v}_{\Ldk}\\
&\lesssim \nor{\beta}_{L^\infty}\sum_{k} \LRp{C_I \nor{J\bm{G}}_{L^\infty\LRp{\Dhat}}  + C_N^2 \nor{J^f}_{L^\infty\LRp{\Dhat}}}\nor{1/J}_{L^\infty\LRp{\Dhat}}\nor{u}_{\Lk}\nor{v}_{\Lk}
\end{align*}
The CFL condition is then determined by mesh-dependent geometric factors of $O(h)$ and order-dependent constants in the inverse and trace inequalities $C_I, C_N$.

@@

For polynomials, $C_I, C_N = O(N^2)$ on the reference element (see, for example \cite{warburton2003constants, ozisik2010constants, chan2015gpu}).

@@

For sufficiently regular solutions, convergence is in $L^2$ as $C (h/N)^{N+1}$ while still maintaining an $O(h/N^2)$ CFL.  Or, if we measure by $h_N = h/N$, CFL is $O(h_N/N)$.  This results from interpolation estimates for one-dimensional B-splines \cite{bazilevs2006isogeometric}

@@

We note that these rules are generated by choosing the number of points $N_q$ such that ${\rm dim}(S_{\rm target}) \leq 2N_q$, and may not be optimal.

\subsection{Reduced quadrature using Gauss-Legendre-Lobatto rules}

Can cast into collocation style using skew-symmetric formulation by using a degree $N + K_{\rm sub} - 1$ Gauss-Legendre or Gauss-Lobatto rule.  Show plots of GLL nodal bases for polynomials vs for B-spline bases - less oscillation for basis functions at nodes on/near the boundary.

>>>>

\newcommand{\Ksub}{K_{\rm sub}}

@@

\date{}
\author{Jesse Chan, John Evans}
\title{Time-domain multi-patch discontinuous Galerkin methods}

@@

\begin{abstract}

Achieve extra $h$-resolution for the same CFL condition by using high order B-splines as a local approximation space on tensor product elements.  Alternatively, achieve an $O(h_N/N)$ CFL with respect to the fine-grid mesh size $h_N = h/N$.

\end{abstract}

@@

Original IGA paper \cite{hughes2005isogeometric}

The ``value of continuity'' for direct solvers \cite{garcia2016value} is to decrease the number of degrees of freedom while maintaining a comparable resolution.

DG on tets using exact geometry \cite{michoski2015foundations}

Multi-patch IGA \cite{langer2015multipatch,langer2016discontinuous}.

%Exploiting tensor product \cite{mantzaflaris2016low}.

Can improve work estimates \cite{chan2015bbdg}, but not CFL.

Improving CFL: Dual-grid filter \cite{warburton2008taming}.  Adding conservation constraints \cite{xu2010conservation,xu2014new}, increasing coupling between elements.  Central DG \cite{reyna2014operator}.

Hermite methods \cite{goodrich2006hermite,vargas2015variations}.

The work here more closely resembles the recently introduced Galerkin difference methods \cite{banks2016galerkin}, which constructs extended-support basis functions from .

\section{Stable timestep restrictions for time-domain DG methods}

Under the method of lines, a PDE is transformed to a system of ordinary differential equations (ODEs) 
\[
\td{\bm{U}}{t} = \bm{A}_h \bm{U}
\]
by discretizing in space.  This system may then be solved using an explicit time stepping, which avoids the inversion of a global matrix.  Since such a scheme requires only matrix-vector products, the storage of a global matrix may also be avoided through a matrix-free implementation.  Under explicit time stepping, the eigenvalues of $\bm{A}_h$ must lie within the stability region of the chosen scheme.

Bounds on real and imaginary parts of eigenvalues can be estimated using theorems of Bendixson and Hirsch \cite{marcus1992survey}, as is done in \cite{chan2015gpu}.

We illustrate this using the advection equation.  
\begin{align*}
V^T A_h U &= \sum_{k} \LRp{\beta\cdot\Grad u,v}_{\Lk} + \LRa{f^*(u),v}_{\Ldk} \\
&\leq \nor{\beta}_{L^\infty}\sum_{k} \nor{\Grad u}\nor{v} + \nor{f^*(u)}_{\Ldk}\nor{v}_{\Ldk}\\
&\lesssim \nor{\beta}_{L^\infty}\sum_{k} \LRp{C_I \nor{J\bm{G}}_{L^\infty\LRp{\Dhat}}  + C_N^2 \nor{J^f}_{L^\infty\LRp{\Dhat}}}\nor{1/J}_{L^\infty\LRp{\Dhat}}\nor{u}_{\Lk}\nor{v}_{\Lk}.  
\end{align*}
The CFL condition is then determined by mesh-dependent geometric factors of $O(h)$ and order-dependent constants in the inverse and trace inequalities $C_I, C_N$.

For shape regular meshes, the ratio between volume and surface area is proportional to $h$, so 
\[
\nor{J^f}_{\LinfDk} \nor{J^{-1}}_{\LinfDk} = O(h).
\]

For polynomials, it can be shown that $C_I,C_N$ both depend quadratically on the degree $N$ (see, for example \cite{warburton2003constants, ozisik2010constants}).  Thus, for DG methods based on locally polynomial approximation spaces, the spectral radius can be bounded by the mesh size, polynomial order, and maximum wavespeed
\[
\rho\LRp{\bm{A}_h} = O\LRp{\frac{\beta_{\max} N^2}{h}}.  
\]
The quadratic dependence of $\rho\LRp{\bm{A}_h}$ on $N$ results in small timesteps if $N$ is large, making the use of high polynomial degrees computationally intractable.

@@

For polynomials, $C_I, C_N = O(N^2)$ on the reference element .

@@

For sufficiently regular solutions, convergence is in $L^2$ as $C (h/N)^{N+1}$ while still maintaining an $O(h/N^2)$ CFL.  Or, if we measure by $h_N = h/N$, CFL is $O(h_N/N)$.  This results from interpolation estimates for one-dimensional B-splines \cite{bazilevs2006isogeometric}.

Show plots of lift functions for polynomials vs for B-spline bases - less oscillation for B-splines.

\begin{figure}
\centering
\subfloat[$N_P = N+\Ksub$]{\includegraphics[width=.4\textwidth]{figs/lift1.png}}
\subfloat[Zoom in]{\includegraphics[width=.4\textwidth]{figs/liftZoom1.png}}\\
\subfloat[$N_P = N$]{\includegraphics[width=.4\textwidth]{figs/lift2.png}}
\subfloat[Zoom in]{\includegraphics[width=.4\textwidth]{figs/liftZoom2.png}}
\caption{Comparison of lift functions for a degree $N_P$ polynomial space and spline space with degree $N = 7$ and $\Ksub = N$ elements.  Degree $N_P$ Gauss-Radau quadrature points are also shown for comparison.  }
\end{figure}

\section{DG formulation}

Skew-symmetric form is energy stable for acoustic, electromagnetic, and elastic wave propagation.  Disadvantage of IGA-DG vs SEM is a lack of equivalence with strong and weak formulations \cite{kopriva2010quadrature}

Penalty fluxes \cite{chan2016short} for dissipation.

Avoid mass matrix inversion using weight-adjusted DG \cite{chan2016weight1,chan2016weight2}.  Especially important for IGA-DG since it decouples the inversion of the mass matrix into a Kronecker structure.

@@

We note that these rules are generated by choosing the number of points $N_q$ such that ${\rm dim}(S_{\rm target}) \leq 2N_q$, and may not be optimal.  The problem is that, to exactly integrate the target space, $N_q = O(N^2)$.  Reduced quadratures are thus necessary.

\subsection{Reduced quadrature rules}

The derivative of a spline does not map into the same spline space; instead, it decreases both the order and continuity.  The derivative operator then satisfies $\pd{}{x} : S^N_{N-1}\rightarrow S^{N-1}_{N-2}$.  To define a derivative operator which maps from $S^N_{N-1}$ to itself, we use an $L^2$ projection.  
\[
\int_{D^k} a(x)\pd{u}{x} v = \int_{\widehat{D}} a(x)\pd{r}{x}J\pd{u}{r} v = \int_{\widehat{D}} a(x)\LRp{\pd{r}{x}J} \Pi_N\LRp{\pd{u}{r}} v.  
\]

Can cast into collocation style using skew-symmetric formulation by using a degree $N + K_{\rm sub} - 1$ Gauss-Legendre or Gauss-Lobatto rule.
